PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$

### 2014-10-21 qfenv_paper/qf_environment.tex
Some edits to bib. Commented out Bundy + Bolzonela

The analysis we describe below uses a fixed cylindrical aperture with
dimensions $R_{\rm{ap}} = 2.9 \;rm{Mpc}$ and $H_{\rm{ap}} = 36
\;\rm{Mpc}$ to measure environment. The same analysis was
extended for varying aperture dimensions $R_{\rm{ap}} = 1.4, 2.9, 4.3 \:
\rm{Mpc}$ and $H_{\rm{ap}} = 36, 71 \;\rm{Mpc}$. We also
adjusted the environment classification thresholds for these
analyses. The results obtained from using different apertures and
environment classification thresholds are qualitatively consistent with the results presented below.

@@

Although we imposed PRIMUS redshift errors on our SDSS galaxies in order to consistently measure environment throughout our entire sample, we note a significant discrepancy between the SDSS and PRIMUS samples in the fractional contribution of each environment classification within the redshift bin. For each of the PRIMUS redshift bins, approximately $50 \%$ of galaxies in the redshift bin are in low density environments and roughly $30 \%$ are in high density environments. In contrast, in the SDSS redshift bin, $38 \%$ of galaxies in the redshift bin are in low density environments and $45 \%$ are in high density environments. We remind the reader that this is due to the different mass-completeness limits imposed on our galaxy sample for each redshift bins. 
\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=0}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments in the redshift range
$z=0$--$0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2.9 \:
\rm{Mpc}$ and $H=36 \: \rm{Mpc}$ and  
classification based on the cut-offs specified in Table \ref{tab:aperture}. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.25$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stellar Mass Function} \label{sec:smf}
Our galaxy sample has so far been classified into quiescent or
star-forming and low or high density environments. We further divide
these subsamples into redshift bins: $0.05-0.12$, $0.2-0.4$,
$0.4-0.6$, and $0.6-0.8$ for a total of 16 subsamples. In Section \ref{sec:smfcalc}, we calculate the SMF for each of these subsamples. Then we examine the evolution of active and quiescent subsample SMFs in different environments in Section \ref{sec:smfevol}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION CALCULATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stellar Mass Function Calculations} \label{sec:smfcalc} 
To calculate the SMFs we employ a non-parametric $1/{V_{\rm{max}}}$ estimator commonly used for galaxy luminosity functions and stellar mass functions, as done in \cite{Moustakas:2013aa} and discussed in the review \cite{Johnston:2011aa}. The differential SMF is given by the following equation:
\begin{equation} \label{eq:phi}
\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log} \:\mathcal{M}) = \sum\limits_{i=1}^{N} \frac{w_i}{V_{\rm{max,avail},i}}. 
\end{equation}
$w_i$ is the statistical weight of galaxy $i$ and $\Phi(\rm{log}\:
\mathcal{M}) \Delta(\rm{log}\: \mathcal{M})$ is the number of galaxies
($N$) per unit volume within the stellar mass range $[\rm{log}
  \:\mathcal{M}, \rm{log} \:\mathcal{M}+\Delta(\rm{log}
  \:\mathcal{M})]$. The equation above is the same as Equation 3 in \cite{Moustakas:2013aa} except that we use $V_{\rm{max,avail}}$ instead than $V_{\rm{max}}$, to account for the edge effects of the survey discussed in Section \ref{sec:edgeeffect}.

@@

While a detailed comparison of the SMFs in each panel for different
epochs is complicated due to the different stellar mass completeness limits, we present some notable trends in each panel. In panel (a), star-forming galaxies in low density environments, we find a significant decrease in the high mass end of the SMF ($\mathcal{M} > 10^{10.75} \mathcal{M}_{\odot}$). In contrast at lower masses ($\mathcal{M} < 10^{10.5} \mathcal{M}_{\odot}$) we observe a notable increase in the SMF. In panel (b), star-forming galaxies in high density environments, we notice a possible decrease above the knee of the SMF ($\mathcal{M} \sim 10^{10.7} \mathcal{M}_{\odot}$) accompanied by an increase below the knee. For the quiescent population in low density environment, shown in panel (c) of Figure \ref{fig:smf}, we notice an increase in $\Phi$ for lower masses, but we do not notice any clear SMF evolution at the massive end. Lastly for the quiescent population in high density environments, panel (d), we find significant increase in $\Phi$ at all masses.

In comparison to results from DEEP2 (\citealt{bundy06a}) and zCOSMOS (\citealt{Bolzonella:2010aa}), some of the noted trends are not immediately evident. However, this is mostly due to the conflicting mass limits proposed in each study. In addition, only part of the redshift range probed by \cite{bundy06a} overlap with our results. Overall, we find that the trends we note in the SMF evolution above are in good agreement with \cite{bundy06a} and \cite{Bolzonella:2010aa}.

@@

\begin{figure*}
    \begin{center}
        \leavevmode
        \epsscale{1.0}
        \epsfig{file=fig_qffit_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_lowenv0_5_highenv3_0_fidmass10_5_lit_primuszerr_baldry_geha_iovino_kovac_peng_envthresh_kovacoffset.eps, height=\textwidth, angle=90}
        \caption{$f_{\rm{Q}}(\mathcal{M}_{\rm{fid}}=10^{10.5} \mathcal{M}_\odot)$ evolution compared to $f_{\rm{red}}(\mathcal{M}_{*}=10^{10.5} \; \mathcal{M}_{\odot})$ in the literature: \cite{Baldry:2006aa} (diamond) and \cite{geha12a} (cross) from SDSS (panel a), \cite{Iovino:2010aa} (empty square) and \cite{Kovac:2014aa} (triangle) from zCOSMOS (panel b), and \cite{Peng:2010aa} (filled square) from SDSS and zCOSMOS (panel c). The $f_{\rm{red}}$ values from \cite{Iovino:2010aa}, \cite{Kovac:2014aa}, \cite{Baldry:2006aa}, and \cite{Peng:2010aa} are calculated from the best-fit parameterizations presented in the respective works. The $f_{\rm{Q}}$ value from \cite{geha12a} is the $f_{\rm{Q}}$ value at $\mathcal{M} = 10^{10.55} \mathcal{M}_{\odot}$. High density environment is represented in red and low density environment is represented in blue. For \cite{Kovac:2014aa} we apply the offset between the color-based and SFR-based galaxy classification in order to plot the $f_{\rm{Q}}$ estimates. We also plot the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}} = 10^{10.5} \mathcal{M}_\odot)$ evolution of our sample with varying environment cut-offs specified on the top right. As in Figure \ref{fig:qffit} the width of the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}} = 10^{10.5} \mathcal{M}_\odot)$ evolution represent the uncertainty in the best-fit parameters of Equation \ref{eq:qffit}.}         \label{fig:qffit_comp}
    \end{center}
\end{figure*}

@@

From our analyses we find the following notable results. The first
three demonstrate that previous findings that are well known in the
local universe are applicable out to $z\sim 0.7$. The last two are
consistent with the findings of \cite{peng10a} but provide increased
detail on the environmental dependence of galaxy evolution:
\begin{enumerate}
% NOT SURE WHAT THIS ONE WAS SAYING:
% \item By measuring the SMF evolution for each of our subsamples, we
%    approximately trace the evolutionary tracks our galaxy population
%    tracing the tracks of mechanisms for ending star formation and
%    galaxies infalling into high density environments.
	\item From the SMFs, we find that the galaxy population in high
    density environments, both star-forming and quiescent, have a
    higher median mass, thus confirming the mass-density relation and
    mass-segregation in different environments throughout our sample's
    redshift range.
	\item For all subsamples, $f_{\rm{Q}}$ increases monotonically with
    galaxy stellar mass, showing a clear mass dependence and
    reflecting the well-established color-mass and morphology-mass
    relations.
	\item We illustrate that $f_{\rm{Q}}$ in high density environments
    is greater than $f_{\rm{Q}}$ in low density environments
    regardless of mass and out to redshift $z\sim 0.7$. This result
    reflects the well known trend that galaxies in high density
    environment are statistically redder, have lower SFRs, and are
    more massive.
	\item $f_{\rm{Q}}$ increases significantly with redshift for both
    low and high density environments. For high density environment,
    this trend is the Butcher-Oemler effect. Furthermore, the
    $f_{\rm{Q}}$ evolution in low density environment suggest the
    existence of internal environment-independent mechanisms for
    ending star formation.
	\item Comparison of the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}})$
    evolution for a range of environment classifications reveals that
    the since $z = 0.8$, $f_{\rm{Q}}$ has evolved by a greater amount
    in the highest density environments. For our purest high
    environment sample ($n_{\rm{env}} > 9$), the total $f_{\rm{Q}}$
    evolution is $\sim 0.1$ greater than the total $f_{\rm{Q}}$
    evolution in low density environment, revealing a moderate
    dependence on environment.
\end{enumerate}

>>>>

The analysis we describe below uses a fixed cylindrical aperture with
dimensions $R_{\rm{ap}} = 2.9 \; \rm{Mpc}$ and $H_{\rm{ap}} = 36
\;\rm{Mpc}$ to measure environment. The same analysis was
extended for varying aperture dimensions $R_{\rm{ap}} = 1.4, 2.9, 4.3 \:
\rm{Mpc}$ and $H_{\rm{ap}} = 36, 71 \;\rm{Mpc}$. We also
adjusted the environment classification thresholds for these
analyses. The results obtained from using different apertures and
environment classification thresholds are qualitatively consistent with the results presented below.

@@

Although we imposed PRIMUS redshift errors on our SDSS galaxies in order to consistently measure environment throughout our entire sample, we note a significant discrepancy between the SDSS and PRIMUS samples in the fractional contribution of each environment classification within the redshift bin. For each of the PRIMUS redshift bins, approximately $50 \%$ of galaxies in the redshift bin are in low density environments and roughly $30 \%$ are in high density environments. In contrast, in the SDSS redshift bin, $38 \%$ of galaxies in the redshift bin are in low density environments and $45 \%$ are in high density environments. We remind the reader that this is due to the different mass-completeness limits imposed on our galaxy sample for each redshift bins. 
\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=0}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments in the redshift range
$z=0$--$0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2.9 \: \rm{Mpc}$ and $H=36 \: \rm{Mpc}$ and  
classified as low environment when $n_{\mathrm{env}} < 0.5$ and as high environment when $n_{\mathrm{env}} > 3.0$. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.2$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stellar Mass Function} \label{sec:smf}
Our galaxy sample has so far been classified into quiescent or
star-forming and low or high density environments. We further divide
these subsamples into redshift bins: $0.05-0.12$, $0.2-0.4$,
$0.4-0.6$, and $0.6-0.8$ for a total of 16 subsamples. In Section \ref{sec:smfcalc}, we calculate the SMF for each of these subsamples. Then we examine the evolution of active and quiescent subsample SMFs in different environments in Section \ref{sec:smfevol}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION CALCULATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stellar Mass Function Calculations} \label{sec:smfcalc} 
To calculate the SMFs we employ a non-parametric $1/{V_{\rm{max}}}$ estimator commonly used for galaxy luminosity functions and stellar mass functions, as done in \cite{Moustakas:2013aa} and discussed in the review \cite{Johnston:2011aa}. The differential SMF is given by the following equation:
\begin{equation} \label{eq:phi}
\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log} \:\mathcal{M}) = \sum\limits_{i=1}^{N} \frac{w_i}{V_{\rm{max,avail},i}}. 
\end{equation}
$w_i$ is the statistical weight of galaxy $i$ and $\Phi(\rm{log}\:
\mathcal{M}) \Delta(\rm{log}\: \mathcal{M})$ is the number of galaxies
($N$) per unit volume within the stellar mass range $[\rm{log}
  \:\mathcal{M}, \rm{log} \:\mathcal{M}+\Delta(\rm{log}
  \:\mathcal{M})]$. The equation above is the same as Equation 3 in \cite{Moustakas:2013aa} except that we use $V_{\rm{max,avail}}$ instead than $V_{\rm{max}}$, to account for the edge effects of the survey discussed in Section \ref{sec:edgeeffect}.

@@

While a detailed comparison of the SMFs in each panel for different
epochs is complicated due to the different stellar mass completeness limits, we present some notable trends in each panel. In panel (a), star-forming galaxies in low density environments, we find a significant decrease in the high mass end of the SMF ($\mathcal{M} > 10^{10.75} \mathcal{M}_{\odot}$). At lower masses ($\mathcal{M} < 10^{10.5} \mathcal{M}_{\odot}$) we observe no noticeable trend in the SMF. In panel (b), star-forming galaxies in high density environments, we notice a possible decrease above the knee of the SMF ($\mathcal{M} \sim 10^{10.7} \mathcal{M}_{\odot}$) accompanied by an increase below the knee. For the quiescent population in low density environment, shown in panel (c) of Figure \ref{fig:smf}, we do not notice any clear SMF evolution throughout the mass range. Lastly for the quiescent population in high density environments, panel (d), we find significant increase in $\Phi$ for lower masses but no notable trend at higher masses.

%In comparison to results from DEEP2 (\citealt{bundy06a}) and zCOSMOS (\citealt{Bolzonella:2010aa}), some of the noted trends are not immediately evident. However, this is mostly due to the conflicting mass limits proposed in each study. In addition, only part of the redshift range probed by \cite{bundy06a} overlap with our results. Overall, we find that the trends we note in the SMF evolution above are in good agreement with \cite{bundy06a} and \cite{Bolzonella:2010aa}.

@@

\begin{figure*}
    \begin{center}
        \leavevmode
        \epsscale{1.0}
        \epsfig{file=fig_qffit_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_lowenv0_5_highenv3_0_fidmass10_5_lit_primuszerr_baldry_geha_iovino_kovac_peng_envthresh_kovacoffset.eps, height=\textwidth, angle=90}
        \caption{$f_{\rm{Q}}(\mathcal{M}_{\rm{fid}}=10^{10.5} \mathcal{M}_\odot)$ evolution compared to $f_{\rm{red}}(\mathcal{M}_{*}=10^{10.5} \; \mathcal{M}_{\odot})$ in the literature: \cite{Baldry:2006aa} (diamond) and \cite{geha12a} (cross) from SDSS (panel a), \cite{Iovino:2010aa} (empty square) and \cite{Kovac:2014aa} (triangle) from zCOSMOS (panel b), and \cite{Peng:2010aa} (filled square) from both SDSS and zCOSMOS (panel c). The $f_{\rm{red}}$ values from \cite{Iovino:2010aa}, \cite{Kovac:2014aa}, \cite{Baldry:2006aa}, and \cite{Peng:2010aa} are calculated from the best-fit parameterizations presented in the respective works. The $f_{\rm{Q}}$ value from \cite{geha12a} is the $f_{\rm{Q}}$ value at $\mathcal{M} = 10^{10.55} \mathcal{M}_{\odot}$. High density environment is represented in red and low density environment is represented in blue. For \cite{Kovac:2014aa} we apply the offset between the color-based and SFR-based galaxy classification in order to plot the $f_{\rm{Q}}$ estimates. We also plot the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}} = 10^{10.5} \mathcal{M}_\odot)$ evolution of our sample with varying environment cut-offs specified on the top right. As in Figure \ref{fig:qffit} the width of the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}} = 10^{10.5} \mathcal{M}_\odot)$ evolution represent the uncertainty in the best-fit parameters of Equation \ref{eq:qffit}.}         \label{fig:qffit_comp}
    \end{center}
\end{figure*}

@@

From our analyses we find the following notable results. The first
three demonstrate that previous findings that are well known in the
local universe are applicable out to $z\sim 0.7$. The last two are
consistent with the findings of \cite{Peng:2010aa} but provide increased
detail on the environmental dependence of galaxy evolution:
\begin{enumerate}
% NOT SURE WHAT THIS ONE WAS SAYING:
% \item By measuring the SMF evolution for each of our subsamples, we
%    approximately trace the evolutionary tracks our galaxy population
%    tracing the tracks of mechanisms for ending star formation and
%    galaxies infalling into high density environments.
	\item From the SMFs, we find that the galaxy population in high
    density environments, both star-forming and quiescent, have a
    higher median mass, thus confirming the mass-density relation and
    mass-segregation in different environments throughout our sample's
    redshift range.
	\item For all subsamples, $f_{\rm{Q}}$ increases monotonically with
    galaxy stellar mass, showing a clear mass dependence and
    reflecting the well-established color-mass and morphology-mass
    relations.
	\item We illustrate that $f_{\rm{Q}}$ in high density environments
    is greater than $f_{\rm{Q}}$ in low density environments
    regardless of mass and out to redshift $z\sim 0.7$. This result
    reflects the well known trend that galaxies in high density
    environment are statistically redder, have lower SFRs, and are
    more massive.
	\item $f_{\rm{Q}}$ increases significantly with redshift for both
    low and high density environments. For high density environment,
    this trend is the Butcher-Oemler effect. Furthermore, the
    $f_{\rm{Q}}$ evolution in low density environment suggest the
    existence of internal environment-independent mechanisms for
    ending star formation.
	\item Comparison of the $f_{\rm{Q}}(\mathcal{M}_{\rm{fid}})$
    evolution for a range of environment classifications reveals that
    the since $z = 0.8$, $f_{\rm{Q}}$ has evolved by a greater amount
    in the highest density environments. For our purest high
    environment sample ($n_{\rm{env}} > 9$), the total $f_{\rm{Q}}$
    evolution is $\sim 0.1$ greater than the total $f_{\rm{Q}}$
    evolution in low density environment, revealing a moderate
    dependence on environment.
\end{enumerate}
